Macroscopic Salt Rejection through Electrostatically Gated Monolayer Porous Graphene

### 2018-10-04 paper/SI.tex
Add Makefiles

\section{Supplementary Section 1.	Graphene membrane manufacturing and 
results.}
\label{sec:exp}

@@

\section{Supplementary Section 2.	Ion diffusion measurement and calculation}
\label{sec:measure}

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$	& 1.95 \\
    Na$^{+}$	& 1.33 \\
    Li$^{+}$	& 1.03 \\
    SO$_{4}^{2-}$	& 1.06 \\
    Ca$^{2+}$	& 0.79\\
    Mg$^{2+}$ &	0.7\\
    Fe(CN)$_{6}^{3-}$	& 0.90\\
    Cl$^{-}$ &	2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivities for the anions and cations of the
    salts used in this work \todo[inline]{do you need any
      reference?}. The theoretical diffusion through PCTE can be
    calculated using Eq. 1 in the main text and the diffusivities of
    the respective anion and cation.}
  \label{tab:diff}
\end{table}

@@

The conductivity increases $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity versus concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl	& 0.0151 $\pm$ 0.00265	& 106	& 1.34 $\pm$ 0.235\\
    NaCl	& 0.0099 $\pm$ 0.000173	& 91	& 1.02 $\pm$ 0.0179\\
    LiCl	& 0.00603 $\pm$ 0.000802	& 83	& 0.679 $\pm$ 0.0903\\
    CaCl$_{2}$	& 0.0125 $\pm$ 0.001	& 149	& 0.789 $\pm$ 0.0631\\
    MgSO$_{4}$	& 0.0088 $\pm$ 0.0012	& 130	& 0.633 $\pm$ 0.084\\
    K$_{3}$Fe(CN)$_{6}$	& 0.024 $\pm$ 0.00275	& 395	& 0.575 $\pm$ 0.0633\\
    K$_{2}$SO4$_{4}$	& 0.0198 $\pm$ 0.00199	& 258	& 0.722 $\pm$ 0.0726\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    $c_{0}$ (mM) & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm)\\
    \hline
    0.1&	0.01903&	30.4\\
    0.33&	0.0541&	16\\
    1	&0.1329	&10\\
    3.3KCl&	0.513&	6\\
    10KCl&	1.08	&3.3\\
    \hline
  \end{tabular}
  \caption{Conductivity and calculated Debye lengths of various KCl
    concentrations in order to obtain rectification versus
    concentration curves. }
  \label{tab:debye}
\end{table}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lcc}
    \hline
    Solution & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm) \\
    \hline
    0.1 mM NaCl	&0.018	&30.4\\
    0.1 mM LiCl	&0.0158	&30.4\\
    0.1 mM CaCl$_{2}$&	0.02187	&17.2\\
    0.1 mM MgSO$_{4}$	&0.0327	&15.2\\
    0.1 mM K$_{2}$SO$_{4}$	&0.0416	&17.2\\
    0.1 mM K$_{3}$(Fe(CN)$_{6}$)&	0.0595	&9.6\\
    \hline
  \end{tabular}
  \caption{conductivity}
  \label{tab:conductivity}
\end{table}

>>>>

\section{Graphene membrane manufacturing and 
results.}
\label{sec:exp}

@@

\section{Ion diffusion measurement and calculation}
\label{sec:measure}

\begin{table}[htbp]
  \centering
  \begin{tabular}{lc}
    \hline
    Salt & $D$ (10$^{9}$ m$^{2}\cdot$s$^{-1}$) \\
    \hline
    K$^{+}$     & 1.95 \\
    Na$^{+}$    & 1.33 \\
    Li$^{+}$    & 1.03 \\
    SO$_{4}^{2-}$       & 1.06 \\
    Ca$^{2+}$   & 0.79\\
    Mg$^{2+}$ & 0.7\\
    {[Fe(CN)$_{6}$]$^{3-}$}   & 0.90\\
    Cl$^{-}$ &  2.03 \\
               \hline
  \end{tabular}
  \caption{Different diffusivities for the anions and cations of the
    salts used in this work \cite{vanysek_2000}. The theoretical diffusion through PCTE can be
    calculated using Eq. 1 in the main text and the diffusivities of
    the respective anion and cation.}
  \label{tab:diff}
\end{table}

@@

The conductivity increases $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity versus concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 0.0151 $\pm$ 0.00265  & 106   & 1.34 $\pm$ 0.235\\
    NaCl        & 0.0099 $\pm$ 0.000173 & 91    & 1.02 $\pm$ 0.0179\\
    LiCl        & 0.00603 $\pm$ 0.000802        & 83    & 0.679 $\pm$ 0.0903\\
    CaCl$_{2}$  & 0.0125 $\pm$ 0.001    & 149   & 0.789 $\pm$ 0.0631\\
    MgSO$_{4}$  & 0.0088 $\pm$ 0.0012   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}$] & 0.024 $\pm$ 0.00275   & 395   & 0.575 $\pm$ 0.0633\\
    K$_{2}$SO4$_{4}$    & 0.0198 $\pm$ 0.00199  & 258   & 0.722 $\pm$ 0.0726\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    $c_{0}$ (mM) & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm)\\
    \hline
    0.1&        0.01903&        30.4\\
    0.33&       0.0541& 16\\
    1   &0.1329 &10\\
    3.3&     0.513&  6\\
    10&      1.08    &3.3\\
    \hline
  \end{tabular}
  \caption{Conductivity and calculated Debye lengths of various KCl
    concentrations in order to obtain rectification versus
    concentration curves. }
  \label{tab:debye}
\end{table}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lcc}
    \hline
    Solution & Conductivity (mS$\cdot$cm$^{-1}$) & $\lambda_{\mathrm{D}}$ (nm) \\
    \hline
    0.1 mM NaCl &0.018  &30.4\\
    0.1 mM LiCl &0.0158 &30.4\\
    0.1 mM CaCl$_{2}$&  0.02187 &17.2\\
    0.1 mM MgSO$_{4}$   &0.0327 &15.2\\
    0.1 mM K$_{2}$SO$_{4}$      &0.0416 &17.2\\
    0.1 mM K$_{3}${[Fe(CN)$_{6}$]}&       0.0595  &9.6\\
    \hline
  \end{tabular}
  \caption{Conductivities and Debye lengths for the 6 different salts
    under investigation.}
  \label{tab:conductivity}
\end{table}

### 2018-10-04 paper/paper.tex
Add Makefiles

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$Fe(CN)$_{6}$,
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity versus time profile
(e.g. \Fig{} \ref{fig:2}a) remains linear within the measurement
duration (1 h), suggesting the concentration change in individual
reservoirs is negligible \todo[inline]{Isn't the increasing of
  conductivity a measure of the diffusion of ions?}. The diffusive fluxes
correspond to different salts were obtained by converting the
conductivity-time profiles with the calibration curves (blue bars in
\Fig{} \ref{fig:2}b and Supplementary \Fig{} S2). For comparison
purposes, based on the assumption that the membrane consists
cylindrical pores with uniform diameter and a low tortuosity $\tau$ of
$\sim{}$1.2 \cite{O_Hern_2012}, the molar diffusive flux through the
bare PCTE membrane, $\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
to those of bare PCTE membrane, suggesting that the mass transfer
resistance of PG is not significant, as the permeability is high. We
observed a more pronounced decrease of diffusive flux in KCl,
K$_{2}$SO$_{4}$ and K$_{3}$Fe(CN)$_{6}$ systems but not for the rest
of the salts. We hypothesized that the strong cation-$\pi$ interactions
between K$^{+}$ cations and graphene slow down the migration of cations
along the surface-mediated transport pathway \cite{Sun_2014}.

@@

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ from at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$Fe(CN)$_{6}$ (details see Supplementary
Section S7 \todo[inline]{Roman, could you check the value of Debye
  length K$3$(Fe(CN)$_{6}$ in Table S3? I calculated 12.4
  nm}). Similar to the KCl system discussed earlier, the salt
rejection was only observed in the positive $V_{\mathrm{G}}$
regime. The degree of salt rejection deceases with the salt valance;
for example, the measured $\xi$ values at $V_{\mathrm{G}}$ = 1.25 V
decreases from $>$ 0.5 for monovalent NaCl to -0.04 for multivalent
K$_{3}$Fe(CN)$_{6}$. This observation is endorsed by our simulations,
in which we calculate $\xi$ as a function of $V_{\mathrm{g}}$ using
the same setups in simulations (\Fig{} \ref{fig:6}b). The calculated
$\xi_{\mathrm{max}}$ values are quantitatively compared with
experiments (\Fig{} \ref{fig:6}c), and the same trend is
observed. Indeed, because the Debye screening length
$\lambda_{\mathrm{D}}$ decreases with the salt valence, the graphene
surface potential appears to decrease in the multivalent salt systems,
thereby reducing $\xi$ (see \Fig{} \ref{fig:5}b). The calculated and
experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d). We notice that without
any fitting parameters, our theory can predict the experimental values
reasonably well. In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of salt diffusivity is
negligible. This finding also explains the concentration dependence of
$\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\delta_{\mathrm{D}} \propto c_{0}^{-0.5}$.

@@

\vspace{1em}
\noindent
\textbf{Membrane preparation}: The ion-diffusion and gating experiments are
performed in a custom-made diffusion cell. An initial dip-wetting
procedure using ethanol/water 1:1 is performed to wet the membrane,
followed by extensive washing and inserting the membrane. The
diffusion cell is filled immediately after mounting the membrane with
DI water to prevent drying. The ion solution is placed on the HCR
side, distilled water on the LCR side, with a volume of 7.33 mL for
each cell. Bar-stirrers close to the membrane ensure well-mixed
solutions avoiding potential external concentration polarization
effects. Water channels in the fixture ensure stable measurement
temperatures using a chiller. The temperature for all measurements is
kept at 25 $^{\circ}$C. A conductivity probe (eDAQ, Pty Ltd.,
Australia), inserted on the LCR side measures the conductivity, which
is recorded with time steps of 2 s. The membrane is kept wet for the
entire series of experiments, multiple rinsing steps with DI water are
performed after each experiment and change to a different salt
solution. The salt solutions were prepared immediately before the
experiments using MilliQ water that was additionally distilled to
increase purity level. The distilled water, has a conductivity of
0.0012 mS/cm with a pH of 6.3, no buffer has been used in these
experiments. 7 salts were used: KCl, NaCl, LiCl, K$_{2}$SO$_{4}$,
CaCl$_{2}$, MgSO$_{4}$ and K$_{3}$Fe(CN)$_{6}$. The salts were
purchased from Sigma Aldrich and used without further purification,
salt solutions were prepared by subsequent dilution of a base solution
(1 M) to the final concentrations.

>>>>

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity versus time profile
(e.g. \Fig{} \ref{fig:2}a) remains linear within the measurement
duration (1 h), suggesting the concentration change in individual
reservoirs is negligible \todo[inline]{Isn't the increasing of
  conductivity a measure of the diffusion of ions?}. The diffusive fluxes
correspond to different salts were obtained by converting the
conductivity-time profiles with the calibration curves (blue bars in
\Fig{} \ref{fig:2}b and Supplementary \Fig{} S2). For comparison
purposes, based on the assumption that the membrane consists
cylindrical pores with uniform diameter and a low tortuosity $\tau$ of
$\sim{}$1.2 \cite{O_Hern_2012}, the molar diffusive flux through the
bare PCTE membrane, $\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
to those of bare PCTE membrane, suggesting that the mass transfer
resistance of PG is not significant, as the permeability is high. We
observed a more pronounced decrease of diffusive flux in KCl,
K$_{2}$SO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$] systems but not for the rest
of the salts. We hypothesized that the strong cation-$\pi$ interactions
between K$^{+}$ cations and graphene slow down the migration of cations
along the surface-mediated transport pathway \cite{Sun_2014}.

@@

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ from at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$[Fe(CN)$_{6}$] (details see Supplementary
Section S7). Similar to the KCl system discussed earlier, the salt
rejection was only observed in the positive $V_{\mathrm{G}}$
regime. The degree of salt rejection deceases with the salt valance;
for example, the measured $\xi$ values at $V_{\mathrm{G}}$ = 1.25 V
decreases from $>$ 0.5 for monovalent NaCl to -0.04 for multivalent
K$_{3}$[Fe(CN)$_{6}$]. This observation is endorsed by our simulations,
in which we calculate $\xi$ as a function of $V_{\mathrm{g}}$ using
the same setups in simulations (\Fig{} \ref{fig:6}b). The calculated
$\xi_{\mathrm{max}}$ values are quantitatively compared with
experiments (\Fig{} \ref{fig:6}c), and the same trend is
observed. Indeed, because the Debye screening length
$\lambda_{\mathrm{D}}$ decreases with the salt valence, the graphene
surface potential appears to decrease in the multivalent salt systems,
thereby reducing $\xi$ (see \Fig{} \ref{fig:5}b). The calculated and
experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d). We notice that without
any fitting parameters, our theory can predict the experimental values
reasonably well. In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of salt diffusivity is
negligible. This finding also explains the concentration dependence of
$\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\lambda_{\mathrm{D}} \propto c_{0}^{-0.5}$.

@@

\vspace{1em}
\noindent
\textbf{Membrane preparation}: The ion-diffusion and gating experiments are
performed in a custom-made diffusion cell. An initial dip-wetting
procedure using ethanol/water 1:1 is performed to wet the membrane,
followed by extensive washing and inserting the membrane. The
diffusion cell is filled immediately after mounting the membrane with
DI water to prevent drying. The ion solution is placed on the HCR
side, distilled water on the LCR side, with a volume of 7.33 mL for
each cell. Bar-stirrers close to the membrane ensure well-mixed
solutions avoiding potential external concentration polarization
effects. Water channels in the fixture ensure stable measurement
temperatures using a chiller. The temperature for all measurements is
kept at 25 $^{\circ}$C. A conductivity probe (eDAQ, Pty Ltd.,
Australia), inserted on the LCR side measures the conductivity, which
is recorded with time steps of 2 s. The membrane is kept wet for the
entire series of experiments, multiple rinsing steps with DI water are
performed after each experiment and change to a different salt
solution. The salt solutions were prepared immediately before the
experiments using MilliQ water that was additionally distilled to
increase purity level. The distilled water, has a conductivity of
0.0012 mS/cm with a pH of 6.3, no buffer has been used in these
experiments. 7 salts were used: KCl, NaCl, LiCl, K$_{2}$SO$_{4}$,
CaCl$_{2}$, MgSO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$]. The salts were
purchased from Sigma Aldrich and used without further purification,
salt solutions were prepared by subsequent dilution of a base solution
(1 M) to the final concentrations.
